---
title: "P8105_HW3"
author: "Meng Fang"
date: '2022-10-6'
output: github_document
---

## Package Loading

```{r}
library(p8105.datasets)
library(tidyverse)
library(dplyr)
library(patchwork)
library(ggridges)
```

## Problem 1

```{r}
data("instacart")
insta <- instacart %>% as_tibble(instacart)
head(insta)
```

#### How many aisles are there, and which aisles are the most items ordered from?

```{r}
insta %>% group_by(aisle) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

We can see that the fresh vegetable aisle is the one that most items ordered from.

#### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
insta %>% group_by(aisle) %>%
  summarise(count = n()) %>%
  filter(count > 10000) %>%
  arrange(desc(count)) %>%
  ggplot(aes(x = reorder(aisle,-count), y = count)) +
  geom_bar(stat = "identity", fill = "#009E73") +
  labs(x = "aisle",
       title = "Number of Items in Each Aisle" ) +
  theme(axis.text.x = element_text(angle = 65, size = 7, hjust = 1), 
        plot.title = element_text(hjust = 0.5))
```

#### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
insta %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(rank = min_rank(-count)) %>%
  filter(rank %in% 1:3) %>%
  knitr::kable()
```

#### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
insta %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  mutate(order_dow = recode(order_dow,
    "0" = "Monday",
    "1" = "Tuesday",
    "2" = "Wednesday",
    "3" = "Thursday",
    "4" = "Friday",
    "5" = "Saturday",
    "6" = "Sunday"
  ) 
  ) %>% pivot_wider(names_from = product_name, values_from = mean_hour) %>%
  as.data.frame()%>%t()%>% knitr::kable()
```

## Problem 2

```{r}
accel <- read_csv(file = "./accel_data.csv") %>% janitor::clean_names()
```

#### Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

- Add a variable indicating whether the day is weekend or weekday.

```{r}
tidy_accel <- accel %>%
  mutate(wke_wkd = ifelse((day == "Saturday"| day =="Sunday"), "weekend", "weekday"),
         day = factor(day, levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
         ) %>%
  pivot_longer(4:1443, names_to = "minute", values_to = "activity_counts") %>%
  mutate(minute = as.integer(sub("activity_","",minute))) %>%
           arrange(week, day)
```

The tidied dataset is has `r nrow(tidy_accel)` rows and `r ncol(tidy_accel)` colomns. The following variables exist in this dataset.

- week
- day_id
- day
- wke_wkd (weekday vs weekend variable)
- minute 
- acitivity_counts

### Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

- Add the total activity together for each day, create variable `activity_total`.

```{r}
tidy_accel %>% group_by(week, day) %>%
  summarize(activity_total = sum(activity_counts)) %>%
  pivot_wider(names_from = day, values_from = activity_total) %>%
  knitr::kable()
```

We don't observe a clear pattern on the table. But we observe two 1440 count on Saturday, which means for every minute the activity count is 1. This is unusual and maybe the patient didn't record data on these two specific day.

#### Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
tidy_accel %>%
  ggplot(aes(x = minute, y = activity_counts, color = day))+
  geom_line(alpha = 0.5)+
  labs(
    title = "24-Hour Activity Counts for Each Day",
    x = "Minute",
    y = "Activity Count"
  ) + 
  theme(plot.title = element_text(hjust = 0.5))
```

We can see a pattern that there are two activity peaks: one around 750 min and one around 1250min in a day.

## Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```

#### Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

```{r}
tidy_noaa <- ny_noaa %>%
  separate(date, into = c("year","month","day"), sep = "-") %>%
  mutate(prcp = as.numeric(prcp), 
         snow = as.numeric(snow), 
         snwd = as.numeric(snwd), 
         tmax = as.numeric(tmax), 
         tmin = as.numeric(tmin),
         month = month.name[as.integer(month)],
         year = as.integer(year))
```

- Find the most commonly observed value for snowfall

```{r}
tidy_noaa %>%
  group_by(snow)%>%
  summarise(count = n())%>%
  arrange(desc(count))
```

Grouping by the snowfall values, we can see that the most frequent observation for snowfall is 0mm.

### Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
tidy_noaa %>%
  filter(month == c("January","July")) %>%
  group_by(month,id, year) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_tmax,color = month)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_wrap(~ month) +
  labs(
    x = "Year",
    y = "Average Max Temperature (tenths of degrees C)",
    title = "Average Max Temperature in January and July"
  )+
  theme(text = element_text(size = 5),
    axis.text.x = element_text(size = 7),
    axis.title = element_text(size = 12),
    strip.text.x = element_text(size = 13),
    plot.title = element_text(size = 14, hjust = 0.5)
    ) 
```

We can see that the average max temperature for each station across years has a wave-like pattern. After some years of increase in average max temperature, we can see a dicrease in the max temperature and then an increase again after some time. For average max temperature in January, there is an apparent outlier in 1982, where the average max temperature is about -110 tens of degrees C. For average max temperature in July, there is an apparent outlier in 1988, where the average max temperature is about 140 tens of degrees C. 

#### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

(i) tmax vs tmin for the full dataset 

```{r}
tmax_tmin <- tidy_noaa %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() +
  scale_fill_continuous(type = "viridis") +
  labs(
    x = "Minimum temperature (tenths of degrees C)",
    y = "Maximum temperature (tenths of degrees C)"
  ) +
  labs(title = "tmax vs tmin") +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5)
  )

sf_dist <- tidy_noaa %>%
  filter(snow > 0 & snow <= 100) %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x = snow, y = year, fill = year)) +
  geom_density_ridges(alpha = 0.5) +
  labs(title = "tmax vs tmin",
       x = "snowfall") + 
  theme(
    axis.text.x = element_text(size = 4),
    axis.text.y = element_text(size = 4),
    axis.text.x.top = element_text(size = 3)
  )

tmax_tmin + sf_dist
```


We can see that for every year, the observed snowfall values follows a similar pattern, where the most observed snowfall value is about 0-40mm. However, we can see an increase in counts of snowfall value between 0-40mm from 2007-2010, which indicates that we may have more snows comparing to previous years.